%-------------------------
% Research CV – Muhammad Hamza
% Clean, professional, ATS-parsable
%-------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{fontawesome5}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage{fontawesome5}
\usepackage[usenames,dvipsnames]{color}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\input{glyphtounicode}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}


% Margins
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Section formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

\pdfgentounicode=1

%-------------------------
% Custom commands

\newcommand{\resumeItem}[1]{
  \item\small{#1 \vspace{-2pt}}
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small #3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeSubSubheading}[2]{
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small #1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeProjectHeading}[2]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \small #1 & #2 \\
    \end{tabular*}\vspace{-5pt}
}

\renewcommand{\labelitemii}{$\circ$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
\begin{document}

%----------HEADING-----------------
\noindent
\begin{tabularx}{\textwidth}{@{} X r @{}}
  \textbf{\Large Muhammad Hamza} &
  \href{mailto:muhammadhamza@kgpian.iitkgp.ac.in}{\faEnvelope\;muhammadhamza@kgpian.iitkgp.ac.in} \\[2pt]
  \href{https://hamza-faarooq.github.io/}{\small hamza-faarooq.github.io} &
  \href{https://github.com/hamza-faarooq}{\faGithub\;hamza-faarooq} \quad
  \href{https://www.linkedin.com/in/muhammad-hamza-iftekhar-hamid-784903204/}{\faLinkedin\;LinkedIn} \\
\end{tabularx}

%-----------EDUCATION-----------------
\section{Education}
\resumeSubHeadingListStart

\resumeSubheading
{Indian Institute of Technology, Kharagpur}{Kharagpur, India}
{Dual Degree (B.Tech \& M.Tech) in Chemical Engineering \;|\; Micro in \textbf{AI \& Applications}}{Aug 2023 -- May 2028}
\resumeItemListStart
\resumeItem {GPA: \textbf{8.62/10}}
\resumeItemListEnd

\resumeSubHeadingListEnd

\vspace{-18pt}

%-----------RESEARCH EXPERIENCE-----------------
\section{Research Experience}
\resumeSubHeadingListStart

%   \resumeSubheading
%     {Indian Institute of Technology, Roorkee \normalfont{(Supervisor: Prof.\ Sparsh Mittal)}}{Remote}
%     {Research Intern – Touch based Emotion Recognition}{May 2026 -- Present}
%   \resumeItemListStart
% \resumeItem{Developed a scalable, robust, end-to-end \textbf{multimodal learning pipeline} integrating textual and visual data, efficiently processing \textbf{54,000+ IMDB reviews} and \textbf{30,000+ movie frames} across \textbf{10+ films} from 4+ genres.}
% \resumeItemListEnd



  \resumeSubheading
    {University of Manchester x NASA \normalfont{(Supervisor: Prof.\ Anirbit Mukherjee)}}{Manchester, UK}
    {Research Intern – Scientific Machine Learning  \href{https://github.com/Hamza-Faarooq/DeepONets}{\faGithub}}{Oct 2025 -- Present}
  \resumeItemListStart
    \resumeItem{Designed and implemented \textbf{DeepONet operators} on spherical domains, modeling heat PDE dynamics with \textbf{maximum harmonic mode  $\mathbf{L_{\max}=511}$}, achieving stable spectral evolution and accurate eigenvalue decay.}
    \resumeItem{Engineered geometry-aware architectures combining \textbf{Graph CNN (HEALPix, $\sim$196K nodes)} \& \textbf{SIREN trunk networks}, mitigating spectral bias and enabling high-frequency function reconstruction on $S^2$ manifolds.}
    \resumeItem{Developed spectral-domain operator learning pipelines using \textbf{spherical harmonic transforms (SHT)}, ensuring alias-free representations up to $\mathbf{L_{\max}=511}$ and preserving orthogonality for robust PDE generalization.}
    % \resumeItem{Conducted \textbf{a priori vs a posteriori validation} of learned operators, demonstrating physically-consistent models improve long-time stability by \textbf{reducing spectral drift \& error accumulation} in iterative time evolution.}
    \resumeItem{Optimized scalable training pipelines via \textbf{stochastic spatial sampling (5K/196K nodes per step)}, significantly reducing the computational cost by \textbf{$\sim$40×} while preserving convergence and operator fidelity.}
    % \resumeItem{Investigated \textbf{spectral operator discovery parallels} between turbulent \textbf{Navier–Stokes closure models and spherical PDEs}, analyzing higher-order differential learning for stable and robust neural operator generalization} 
  \resumeItemListEnd

  \resumeSubheading
    {Indian Institute of Management, Ranchi \normalfont{(Supervisor: Prof.\ Sobhan Sarkar)}}{Remote}
    {Research Intern – Multimodal Machine Learning \href{https://github.com/Hamza-Faarooq/IIM_Research_Intern}{\faGithub}}{April 2025 -- Aug 2025}
  \resumeItemListStart
\resumeItem{Developed a scalable, robust, end-to-end multimodal learning pipeline, integrating textual \& visual data, efficiently processing \textbf{54,000+ IMDB reviews} and \textbf{30,000+ movie frames} across \textbf{10+ movies} from \textbf{4+ genres.}}

\resumeItem{Constructed \& benchmarked architectures combining \textbf{BERT embeddings}, \textbf{Vision Transformer (ViT)} features, and \textbf{CLIP-based alignment} under multiple fusion strategies for improved multimodal performance evaluation.}  

\resumeItem{Evaluated the models on \textbf{five rigorous performance metrics} (Accuracy, F1, AUC, MAP, NDCG), achieving exceptionally strong peak performance of \textbf{0.875 Accuracy} and \textbf{0.982 NDCG}, best on \textbf{'Interstellar'} dataset.}
\resumeItem{Introduced the \textbf{Dempster--Shafer evidential fusion} to explicitly model cross-modal uncertainty \& belief conflict, enabling adaptive, confidence-weighted multimodal predictions improving robustness over baseline fusion methods.}
% \resumeItem{Conducted a \textbf{comprehensive comparative analysis of multimodal fusion strategies}, evaluating the \textbf{concatenation, attention-based, and evidential approaches} to model cross-modal interactions, significantly improving ranking performance, semantic alignment, and \textbf{uncertainty-aware decision-making robustness}}
\resumeItemListEnd


\vspace{-3pt}

  \resumeSubheading
    {University of Oulu, Finland  \normalfont{(Supervisor: Prof.\ Shuai Lee)}}{Remote}
    {Research Intern – Constrained Quadratic Optimisation \href{https://github.com/Hamza-Faarooq/Pruning-with-SNNs}{\faGithub}}{May 2025 -- June 2025}
  \resumeItemListStart
\resumeItem{Formulated \& solved constrained quadratic programs for pruning Spiking Neural Networks (SNNs), achieving \textbf{87.5\% spatial sparsity} while preserving accuracy, \textbf{using CVXPY} with biologically inspired constraints.}
\resumeItem{Integrated criticality-based neuron selection into pruning pipeline using layer-wise importance scores, \textbf{retaining top 112 out of 784 neurons}, and visualizing accuracy versus criticality thresholds to guide structured pruning.}
    \resumeItem{Implemented \textbf{KL-divergence-based temporal sparsity regularization}, penalizing redundant spiking across timesteps and integrating it as a weighted constraint in CVXPY to jointly reduce spatial and temporal redundancy.}
    % \resumeItem{\textbf{Conducted experiments on MNIST and CIFAR-10 with rate encoding}, benchmarking the \textbf{accuracy degradation under varying sparsity levels} and \textbf{comparing  the SNN pruning to the ANN baselines.}}
  \resumeItemListEnd

\resumeSubHeadingListEnd

\vspace{-16pt}

%-----------PUBLICATIONS-----------
\section{Publications \& Patents}
\vspace{-1pt}

\noindent\begin{tabularx}{\textwidth}{@{}l X@{}}
\small\textit{Legend:} &
\small C\,=\,Conference, \; J\,=\,Journal, \; P\,=\,Preprint, \; U\,=\,Under Review \\[6pt]

\textbf{[J.1, C]} &
\textbf{ Muhammad Hamza et al.} (2025).
\href{https://ieeexplore.ieee.org/abstract/document/11492803}{Movie Review Helpfulness Prediction using Semantic Alignment \& Evidential Fusion-based Multimodal Approach}.
\textbf{\textit{ICEEE--IEEE Conference 2026}}. \\
[16pt]

% \textbf{[J.2, U]} &
% \textbf{Muhammad Hamza}.
% \textit{Criticality-Constrained Iterative Pruning for Energy-Efficient Spiking Neural Networks via Combined Importance Scoring}.
% \textbf{Under Review}. \\
% [22pt]

\textbf{[J.2, P]} &
\textbf{ Muhammad Hamza}. (2026).
\href{https://arxiv.org/abs/2606.30455}{\textit{Curvature-Weighted Gradient Diversity: A Noise Measure for Geometry-Adaptive SGD Schedules}}.
\textbf{arXiv:2606.30455}. \\

\end{tabularx}

\vspace{-8pt}

%-----------PROJECTS-----------------
\section{Projects}
\resumeSubHeadingListStart


\resumeProjectHeading
{\textbf{ApexSearch: F1 Audio-Visual RAG \& Multimodal Retrieval Engine}
\href{https://github.com/Hamza-Faarooq/ApexSearch}{\faGithub}
$\mid$ \emph{PyTorch, FAISS, CLIP, CLAP}}{2026}

\resumeItemListStart

\resumeItem{Indexed \textbf{2000+ 1-FPS} video frames and \textbf{1-second audio segments} across 4 distinct F1 broadcast events, building a \textbf{cross-modal FAISS index} supporting joint audio-visual similarity queries at runtime.}
\resumeItem{Engineered a multimodal RAG pipeline for Formula-I broadcasts, embedding \textbf{1FPS video frames \& 1-second audio segments} using \textbf{CLIP \& CLAP into FAISS}, enabling \textbf{sub-50ms retrieval} across acoustic race events.}
\resumeItem{Integrated a hosted \textbf{LLaVA-based} vision-language model through the \textbf{Groq API} to generate Formula-I commentary from multimodal evidence, eliminating GPU requirements while enabling Streamlit deployment.}

\resumeItemListEnd

\resumeProjectHeading
{\textbf{VictorianGPT: Compound RAG Architecture} \href{https://github.com/Hamza-Faarooq/VictorianGPT}{\faGithub} \textbar\ \emph{PyTorch, QLoRA, ChromaDB, HuggingFace}}{2026}

\resumeItemListStart

\resumeItem{Built a domain-specific instruction-tuning dataset of \textbf{20K+ Victorian dialogue pairs} via regex extraction and quality filtering across \textbf{15 Project Gutenberg novels}.}

\resumeItem{Fine-tuned a \textbf{3B-parameter LLM with QLoRA} 4-bit quantization, optimizing efficiency \& training stability.}

\resumeItem{Deployed a \textbf{ChromaDB-backed RAG system} on Hugging Face, implementing retrieval weighting to balance retrieved context and generation quality.}

\resumeItemListEnd

\resumeProjectHeading
  {\textbf{Visuo-Lingual Affective Recommender System (VLARS)} \href{https://github.com/Hamza-Faarooq/Visuo-Lingual-Affective-Recommender-System-VLARS}{\faGithub} \textbar\ \emph{PyTorch, FAISS, FastAPI, Docker, GCP}}{2026}
\resumeItemListStart
  
  \resumeItem{Built a multimodal user-state encoder fusing \textbf{DeepFace} micro-expression vectors and \textbf{DistilRoBERTa} sentiment embeddings into \textbf{a unified 512-dim} affective representation for downstream retrieval.}
  
  \resumeItem{Designed an L1-normalized late-fusion heuristic resolving cross-modal contradictions, \textbf{improving recall@5 by 23\% over text-only and 18\% over face-only} baselines on a 500-clip eval set.}
  
  \resumeItem{Implemented high-dimensional semantic search across 5,000 movies using \textbf{MiniLM embeddings \& FAISS retrieval}, achieving an average end-to-end \textbf{inference latency of 180ms} on CPU.}
    
\resumeItemListEnd

  
% \resumeProjectHeading
%   {\textbf{Visuo-Lingual Affective Recommender System (VLARS)} \href{https://github.com/Hamza-Faarooq/Visuo-Lingual-Affective-Recommender-System-VLARS}{\faGithub} \textbar\ \emph{PyTorch, FAISS, DeepFace, Transformers}}{2026}
% \resumeItemListStart
  
%   \resumeItem{Engineered a multimodal recommendation engine \textbf{fusing facial micro-expressions (DeepFace) and textual sentiment (DistilRoBERTa)} pipelines to synthesize physiological and psychological user states.}
  
%   \resumeItem{Designed an \textbf{$L_1$-normalized late-fusion heuristic} resolving cross-modal contradictions, improving\textbf{ recall@5 by 23\%} over text-only and \textbf{18\%} over face-only baselines on a \textbf{500-clip eval set.}}
  
%   \resumeItem{Implemented high-dimensional semantic search across 5,000 movies using \textbf{MiniLM embeddings \& FAISS retrieval}, achieving an \textbf{average end-to-end inference latency of 180ms on CPU.}}
  




\resumeSubHeadingListEnd

%-----------SCHOLASTIC ACHIEVEMENTS-----------
\section{Scholastic Achievements}
\resumeSubHeadingListStart
  \resumeItemListStart
    \resumeItem{Ranked in the \textbf{top 0.26\%} of \textbf{1.2M+ candidates} in the Joint Entrance Examination \textbf{(JEE) Main}, 2023.}
    \vspace{-2pt}
    \resumeItem{Ranked in the \textbf{top 3\%} among \textbf{200K+ candidates} in Joint Entrance Examination \textbf{(JEE) Advanced}, 2023.}
  \resumeItemListEnd
\resumeSubHeadingListEnd

\vspace{-10pt}
%-----------COMPETITIONS-----------
\section{Competitions}
\resumeSubHeadingListStart

\vspace{2pt}
  \resumeSubheading
    {Optiver Trade-athon 2026}{\textbf{Optiver}}{March 2026}{}
  \resumeItemListStart
    \resumeItem{\textbf{Rank 23} in \textbf{Optiver's Trade-athon}, a trading challenge involving real-time market-making \& pricing strategies}
  \resumeItemListEnd

%   \resumeSubheading
%     {Convolve 4.0 -- Pan-IIT AI/ML Hackathon}{\textbf{Quadeye Securities/IIT Madras}}{Feb 2026}{}
%   \resumeItemListStart
% \resumeItem{Secured \textbf{All India Rank 102} in the highly competitive screening ML/AI assessment amongst \textbf{3,000+ students.}}
% \resumeItemListEnd

  \resumeSubheading
    {NK Securities Research Hackathon 2025}{\textbf{NK Securities Research}}{Jun 2025}{}
  \resumeItemListStart
    \resumeItem{Secured \textbf{Global Rank 198/6000+} and achieved an ultra-low MSE of \textbf{9.8e-5} in implied volatility prediction using advanced feature engineering, \textbf{outperforming standard ML baselines by 10$\times$ in accuracy.}}
    \resumeItem{Developed a high-frequency volatility forecasting pipeline using \textbf{278K+ records} and \textbf{90+ engineered features}.}
  \resumeItemListEnd

  \resumeSubheading
    {Convolve 3.0 -- Pan-IIT AI/ML Hackathon}{\textbf{IIT Guwahati}}{Jan 2025}{}
  \resumeItemListStart
    \resumeItem{Secured \textbf{Top 50 among 4,000+ teams} by developing a data-driven framework for loan default prediction under real-world constraints, demonstrating robust generalization and decision-aware modeling.}
    \resumeItem{Designed an \textbf{XGBoost-based predictive model} with robust preprocessing for missing data and severe class imbalance, achieving \textbf{0.85 precision} and \textbf{0.91 F1-score} for reliable risk-sensitive predictions.}
  \resumeItemListEnd

\resumeSubHeadingListEnd

%-----------TECHNICAL SKILLS-----------
\section{Technical Skills}

\small
\textbf{Languages \& Frameworks:} Python, C++, C, SQL, PyTorch, TensorFlow, NumPy, Pandas, Scikit-learn, SciPy \\[2pt]
\textbf{ML / AI:} Deep Learning, Neural Operators (DeepONet, Graph-DeepONet), Transformers (ViT, BERT, CLIP), NLP, Computer Vision, Multimodal Learning, Spiking Neural Networks (SNNs) \\[2pt]
\textbf{Tools:} Git, \LaTeX


\resumeSubHeadingListEnd


\section{Coursework Information}
\vspace{2pt}
\begin{itemize}[leftmargin=0.05in, label={}]
\item \textbf{Mathematics and Computer Science:} Artificial Intelligence: Foundations and Applications, Programming and Data Structures (Theory and Laboratory), Linear Algebra, Advanced Calculus, Transform Calculus, Numerical and Complex Analysis, Computer-Aided Process Engineering
\vspace{-5pt}
\item \textbf{MOOCs:} Classical Machine Learning, Neural Networks and Deep Learning, Improving Deep Neural Networks, Convolutional Neural Networks, Natural Language Processing, Large Language Models
\end{itemize}

%-------------------------------------------
\end{document}
